Columnar VCF for Biobank scale datasets using Zarr

### 2024-04-11 paper.tex
More plot infrastructure for column extract

@@

>>>>

\begin{figure}
\includegraphics{figures/column-extract}
\caption{Time to extract the genome position and write to a text file.
Total CPU time required to extract the POS field for BCF,
sav and Zar formats
for the data in Figure~\ref{fig-data-storage}.
For the BCF file we used \texttt{bcftools query -f"\%POS$\backslash$n"}.
For sav, we used the Savvy C++ API to extract just the position, 
and output text using the \texttt{std::cout} stream. For Zarr, we read 
\texttt{variant\_position} into a NumPy array, and then wrote to
a text file using the Pandas
\texttt{write\_csv} method. Writing
text dominates in the case of Zarr: populating the NumPy array 
in memory takes less than a second in all cases. Genozip does
not offer an option to extract individual columns, and so would require
the output to be piped into \texttt{bcftools query} or a
Unix utility such as \texttt{cut}.
\label{fig-column-extract}}
\end{figure}

\subsection{Extracting fields}
We have focused on the genotype matrix up to this point, contrasting
Zarr with existing row-wise methods. The genotype matrix, however,
is only one aspect of VCF data.
Real-world VCFs encapsulate much more than just the genotype
matrix, and can contain large numbers of additional fields. Efficient
extraction of the data from these fields is a key use-case
and an area where columnar storage has clear benefits.
Figure~\ref{fig-column-extract} shows the time required to extract 
the genomic position of each variant in the simulated benchmark 
dataset, which we can use as an indicative example of a per-variant 
query. Although Savvy is many times faster than \texttt{bcftools query}
here, the row-wise storage strategy that they share means that 
the entire dataset must be read into memory and 
decompressed to extract just one field from each record. Zarr's
columnar storage excels at these tasks: we only read and decompress
the information required.

@@

Reel off Python data science ecosystem: Pandas~\citep{mckinney2010data} etc
